---
title: "Using evacpath in a new region"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using evacpath in a new region}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE, purl = FALSE)
checklist <- data.frame(
  item = c(
    "Hazard or inundation input", "Road or pathway input", "DEM or elevation input",
    "Target CRS", "Walking speed", "Local exclusions", "Output directory", "QA/QC review"
  ),
  review = c(
    "A defined hazard footprint and its intended threshold.",
    "A connected, suitable pedestrian network with documented attributes.",
    "Coverage across the study area, elevation sign convention, and resolution.",
    "A projected CRS in meters for distance calculations.",
    "A documented assumption appropriate to the planning scenario.",
    "Piers, private roads, closures, or other locally unsuitable features.",
    "An explicit directory when using write_evac_outputs().",
    "CRS, geometry, extent, origins, exits, and reachability checks."
  )
)
```

For a new region, `evacpath` expects three core spatial inputs:

1. a hazard or inundation zone,
2. a road/pathway network, and
3. a DEM or elevation raster.

The most important requirement is that distance calculations should use a projected coordinate reference system in meters.

The workflow builds on open-source least-cost path methods for evacuation planning (Cordero et al. 2025) and uses `leastcostpath` for least-cost path and movement-potential modeling (Lewis 2023; Lewis 2021).

## Regional checklist

```{r checklist, echo = FALSE, eval = TRUE}
knitr::kable(checklist, col.names = c("Item", "Review before modeling"))
```

```{r setup}
library(evacpath)
```

## Minimal workflow

```{r minimal}
result <- run_evacpath(
  hazard_zone = "path/to/hazard_zone.gpkg",
  roads = "path/to/roads.gpkg",
  dem = "path/to/dem.tif",
  target_crs = "EPSG:XXXX",
  region_name = "New region",
  walking_speed_mps = 1.22
)
```

## Common customizations

```{r custom}
result <- run_evacpath(
  hazard_zone = "path/to/hazard_zone.gpkg",
  roads = "path/to/roads.gpkg",
  dem = "path/to/dem.tif",
  target_crs = "EPSG:XXXX",
  region_name = "New region",

  # Road and escape-point geometry assumptions
  road_buffer_m = 2,
  escape_buffer_m = 5,
  final_road_buffer_m = 3,

  # Origin-point density
  grid_resolution = 50,
  max_origins = NULL,

  # Travel-time assumption
  walking_speed_mps = 1.22,

  # Output clipping
  clip_mode = "road_hazard"
)
```

## Case-specific preprocessing

Keep region-specific choices outside the package. For example, if your DEM uses negative elevation for land, correct it before calling `run_evacpath()`. If your road layer contains piers or other unsuitable features, remove them with `clean_roads()`.

```{r preprocessing}
roads <- clean_roads(
  roads = "path/to/roads.gpkg",
  exclude = list(field = "road_type", values = c("pier", "private")),
  target_crs = "EPSG:XXXX"
)
```

## References

Cordero, E., Ruiz Vélez, R., Huérfano Moreno, V., Sherman, C., 2025. Enhancing tsunami evacuation strategies in Puerto Rico using open-source least-cost path analysis. J. Disaster Sci. Manag. 1, 18. https://doi.org/10.1007/s44367-025-00018-y

Lewis, J., 2023. leastcostpath: Modelling Pathways and Movement Potential Within a Landscape.

Lewis, J., 2021. Probabilistic Modelling for Incorporating Uncertainty in Least Cost Path Results: a Postdictive Roman Road Case Study. Journal of Archaeological Method and Theory 28, 911-924. https://doi.org/10.1007/s10816-021-09522-w
