---
title: "Modeling assumptions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modeling assumptions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

Evacuation maps are conditional on the spatial inputs and modeling choices that
produce them. Record those choices and inspect their effects before using a map
for planning.

## Coordinate reference system

Use a projected CRS in meters. Buffers, grid resolution, route distance, and
walking-speed conversion are not meaningful in longitude and latitude degrees.

## Hazard, terrain, and road inputs

- Define the hazard zone for the question being asked; for tsunami modeling,
  distinguish the land-only origin zone from the escape boundary.
- Confirm the digital elevation model sign convention before using land and
  water thresholds. Do not assume every source uses positive values for land.
- Review road attributes and remove unsuitable features such as piers or routes
  that are not usable by pedestrians.

## Buffers and origin density

`road_buffer_m` controls how far movement can extend from the road centerline.
`escape_buffer_m` provides local access around candidate escape points.
`grid_resolution` controls the sampling density of origins; finer grids increase
runtime and should be justified by the terrain and road data resolution.

## Conductance and least-cost paths

`make_conductance_surface()` uses a slope-based conductance surface from
`leastcostpath`. The following arguments affect route geometry rather than just
the final time conversion:

```{r lcp}
run_evacpath(
  ...,
  lcp_cost_function = "tobler",
  lcp_neighbours = 16,
  lcp_crit_slope = 12,
  lcp_max_slope = NULL
)
```

`walking_speed_mps` converts modeled route distance into time. It should be set
to the population and planning scenario of interest.

## Output clipping

`clip_mode = "hazard"` maps the time surface across the full hazard zone.
`"road_hazard"` limits output to the road-buffer area, `"hazard_plus_roads"`
retains both, and `"none"` leaves the Voronoi output unclipped. Choose the mode
that matches the intended map interpretation.
