---
title: "Scenario comparison"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scenario comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`compare_evac_scenarios()` runs the same study inputs under named parameter
overrides. It returns each raw result, a scenario summary, and combined spatial
layers when available.

```{r compare}
comparison <- compare_evac_scenarios(
  hazard_zone = hazard_zone,
  roads = roads,
  dem = dem,
  target_crs = "EPSG:32748",
  max_origins = 100,
  max_destinations = 25,
  scenarios = list(
    baseline = list(walking_speed_mps = 1.22),
    slow_walkers = list(walking_speed_mps = 0.75),
    conservative_lcp = list(
      lcp_neighbours = 8,
      lcp_max_slope = 30
    )
  )
)

comparison$summary
```

The summary includes origin and destination counts, time statistics, threshold
percentages, walking speed, and the selected least-cost settings. A scenario
with incomplete outputs is retained with missing summary values rather than
discarding the entire comparison.

Treat scenario outputs as sensitivity analysis. A slower walking speed changes
time conversion; conductance and neighbourhood settings can change modeled route
geometry as well as distance.
